---
title: "Project 1_Hao Hu"

output: html_notebook
---

# Inferring reasons why a huge reduction between people who vote for Democrat in 2012 and 2016 exist

```{r}
getwd()
setwd("C:/Users/sluo1/Desktop/5243/Spring2021-Project1-hh2874/output")


```

```{r echo=FALSE, warning = FALSE}
install.packages("tidyverse")
install.packages("DT")
library(rbenchmark)
library(readr)
library(DT)
library(tibble)
library(ggplot2)
library(haven)
library(dplyr)
```

```{r echo=FALSE}
anes_dat <-read_sav("C:/Users/sluo1/Desktop/5243/Spring2021-Project1-hh2874/data/anes_timeseries_cdf_sav/anes_timeseries_cdf.sav")


```



```{r echo=FALSE}
save(anes_dat, file = "anes_dat")


```
 
```{r echo=FALSE, warning = FALSE}
Election_years=as.character(c(2012,2016))

anes_use=anes_dat%>%
  mutate(
    year=as_factor(VCF0004),
    turnout=as_factor(VCF0703),
    vote=as_factor(VCF0706),
    race=as_factor(VCF0105a),
    gender=as_factor(VCF0104),
    working_status=as_factor(VCF0116),
    social_class=as_factor(VCF0148),
    Difference=as_factor(VCF0501),
    religion=as_factor(VCF0128),
    church_att=as_factor(VCF0130),
    religion_important=as_factor(VCF0846)
)%>%
filter(year %in% Election_years)

library(data.table)
Table1<-data.table(anes_use%>%
             select(year, turnout, vote, race, gender,working_status,social_class,Difference)%>%
             filter(!is.na(turnout)))%>%filter(year=="2016" | year=="2012")
```

As the American election ended a while ago, Joe Biden was inaugurated as the 46th president of the United States. United States election is one of the biggest event in every people in the United States. That makes people think a lot about what affect the voting result.    

The line chart below shows the voting result in ANES data. From the chart, People who vote for Democrat party is much greater than people who vote for Republican party in 2012 and in 2016 People who vote for Democrat party is still greater than Republican party but it is not that too much. What causes that happened? A huge reduction in the number of voters who vote for Democrat might be caused by many reasons.          

```{r echo=FALSE}
anes_count<-Table1 %>% filter(turnout=="3. Voted (registered)")
anes_count_Democrat_12<-nrow(anes_count %>% filter(vote=="1. Democrat") %>% filter(year=="2012"))
anes_count_Republican_12<-nrow(anes_count %>% filter(vote=="2. Republican") %>% filter(year=="2012"))

anes_count_Democrat_16<-nrow(anes_count %>% filter(vote=="1. Democrat") %>% filter(year=="2016"))
anes_count_Republican_16<-nrow(anes_count %>% filter(vote=="2. Republican") %>% filter(year=="2016"))

table_party<-data.frame(year=rep(c("2012","2016"),each=2),
           party=rep(c("Democrat","Republican"),2),
           count=c(anes_count_Democrat_12,anes_count_Republican_12,anes_count_Democrat_16,anes_count_Republican_16))

ggplot(data=table_party,aes(x=year,y=count,group=party))+
  geom_line(aes(color=party))+
  geom_point()+ggtitle("Voting condition in 2012 and 2016")


```




There are so many factors may affect the voting. Now, focusing on Religious beliefs. Does what voter believe might affect the voting result? According to our analysis of ANES data, religious beliefs affect the voting result indeed.    


 


First up, as the bar chart below shows, people think religious beliefs is important in voting which is twice as much as who think is not important in both 2012 and 2016. From this chart, we can preliminary conclude that religious beliefs have some impact on voting.       
```{r echo=FALSE}
anes_religion_important=anes_dat%>%
  mutate(
    year=as_factor(VCF0004),
    vote=as_factor(VCF0706),
    religion_important=as_factor(VCF0846)
)%>%
filter(year %in% Election_years)%>%
  select(year,vote,religion_important)%>%filter(!is.na(vote))%>%filter(!is.na(religion_important))

Table_religion_important_2012<-anes_religion_important[anes_religion_important$year=="2012",]
Table_religion_important_2016<-anes_religion_important[anes_religion_important$year=="2016",]

Table_religion_important_2012$yes_count<-ifelse(Table_religion_important_2012$religion_important=="1. Yes, important",1,0)
Table_religion_important_2012$no_count<-ifelse(Table_religion_important_2012$religion_important=="2. No, not important",1,0)

Table_religion_important_2016$yes_count<-ifelse(Table_religion_important_2016$religion_important=="1. Yes, important",1,0)
Table_religion_important_2016$no_count<-ifelse(Table_religion_important_2016$religion_important=="2. No, not important",1,0)

num_countyes2012<-sum(Table_religion_important_2012$yes_count)
num_countno2012<-sum(Table_religion_important_2012$no_count)
num_countyes2016<-sum(Table_religion_important_2016$yes_count)
num_countno2016<-sum(Table_religion_important_2016$no_count)

religion_important<-data.frame(year=rep(c("2012","2016"),each=2),
               attitude=rep(c("yes","no"),2),
               count=c(num_countyes2012,num_countno2012,num_countyes2016,num_countno2016))

ggplot(data=religion_important, aes(x=year, y=count, fill=attitude)) +
  geom_bar(stat="identity", position="dodge2")+
  ggtitle("People think religion is important or not")+
  geom_text(aes(label=count), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  scale_fill_brewer(palette="Accent")+theme_minimal()





```



### Protestant has much proportion on Republican in 2016.    

```{r echo=FALSE}
anes_religion=anes_dat%>%
  mutate(
    year=as_factor(VCF0004),
    vote=as_factor(VCF0706),
    religion=as_factor(VCF0128),
    church_att=as_factor(VCF0130)
)%>%
filter(year %in% Election_years)%>%
  select(year,vote,religion,church_att)%>%filter(!is.na(vote))%>%filter(!is.na(religion))%>%filter(!is.na(church_att))
```


But how believes affect the religious. The pie charts show the distribution of voters in different religions voting condition. In 2012, it is obvious that Protestant has weighted approximately 50 percent of the total proportion on both Democrat and Republican and Jewish only weighted a tiny part of the total proportion on both Democrat and Republican. Unfortunately, it is hard for us to find out how beliefs affect the voting result in 2012.          
```{r echo=FALSE}
#2012 analysis

DP2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="1. Democrat")%>%filter(religion=="1. Protestant"))
DC2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="1. Democrat")%>%filter(religion=="2. Catholic [Roman Catholic]"))
DJ2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="1. Democrat")%>%filter(religion=="3. Jewish"))
DO2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="1. Democrat")%>%filter(religion=="4. Other and none (also includes DK preference)"))

RP2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="2. Republican")%>%filter(religion=="1. Protestant"))
RC2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="2. Republican")%>%filter(religion=="2. Catholic [Roman Catholic]"))
RJ2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="2. Republican")%>%filter(religion=="3. Jewish"))
RO2012<-nrow(anes_religion%>%filter(year=="2012")%>%filter(vote=="2. Republican")%>%filter(religion=="4. Other and none (also includes DK preference)"))

sum<-sum(DP2012,DC2012,DJ2012,DO2012)
table_religion_D2012<-data.frame(religion=c("Protestant","Other and none","Catholic","Jewish"),
                           count=c(DP2012,DO2012,DC2012,DJ2012),
                           prop=c(round(DP2012/sum,4)*100,round(DO2012/sum,4)*100,round(DC2012/sum,4)*100,round(DJ2012/sum,4)*100))

sum<-sum(RP2012,RC2012,RJ2012,RO2012)
table_religion_R2012<-data.frame(religion=c("Protestant","Other and none","Catholic","Jewish"),
                           count=c(RP2012,RO2012,RC2012,RJ2012),
                           prop=c(round(RP2012/sum,4)*100,round(RO2012/sum,4)*100,round(RC2012/sum,4)*100,round(RJ2012/sum,4)*100))

table_religion_D2012 <- table_religion_D2012 %>%
  arrange(desc(religion)) %>%
  mutate(lab.ypos1 = cumsum(prop) - 0.5*prop)

table_religion_R2012 <- table_religion_R2012 %>%
  arrange(desc(religion)) %>%
  mutate(lab.ypos2 = cumsum(prop) - 0.5*prop)


ggplot(table_religion_D2012, aes(x ="" , y = prop, fill = religion))+
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start = 0)+
  geom_text(aes(y = lab.ypos1, label = paste(prop,"%", sep = "")), col = "white")+ggtitle("2012 diff religions vote for Democrat")
  
ggplot(table_religion_R2012, aes(x ="" , y = prop, fill = religion))+
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start = 0)+
  geom_text(aes(y = lab.ypos2, label = paste(prop,"%", sep = "")), col = "white")+ggtitle("2012 diff religions vote for Republican")

```

In 2016 pie chart, it is quite different from the pie chart in 2012 since the proportion of some religions change a lot. It seems that lots of Protestants voted for Republican in 2016 since the proportion of the Protestants decrease a lot comparing to 2012. That might explain why there exist a huge gap in the line chart above.       

```{r echo=FALSE}
#2016 analysis

DP2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="1. Democrat")%>%filter(religion=="1. Protestant"))
DC2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="1. Democrat")%>%filter(religion=="2. Catholic [Roman Catholic]"))
DJ2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="1. Democrat")%>%filter(religion=="3. Jewish"))
DO2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="1. Democrat")%>%filter(religion=="4. Other and none (also includes DK preference)"))

RP2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="2. Republican")%>%filter(religion=="1. Protestant"))
RC2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="2. Republican")%>%filter(religion=="2. Catholic [Roman Catholic]"))
RJ2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="2. Republican")%>%filter(religion=="3. Jewish"))
RO2016<-nrow(anes_religion%>%filter(year=="2016")%>%filter(vote=="2. Republican")%>%filter(religion=="4. Other and none (also includes DK preference)"))

sum<-sum(DP2016,DC2016,DJ2016,DO2016)
table_religion_D2016<-data.frame(religion=c("Protestant","Other and none","Catholic","Jewish"),
                           count=c(DP2016,DO2016,DC2016,DJ2016),
                           prop=c(round(DP2016/sum,4)*100,round(DO2016/sum,4)*100,round(DC2016/sum,4)*100,round(DJ2016/sum,4)*100))

sum<-sum(RP2016,RC2016,RJ2016,RO2016)
table_religion_R2016<-data.frame(religion=c("Protestant","Other and none","Catholic","Jewish"),
                           count=c(RP2016,RO2016,RC2016,RJ2016),
                           prop=c(round(RP2016/sum,4)*100,round(RO2016/sum,4)*100,round(RC2016/sum,4)*100,round(RJ2016/sum,4)*100))

table_religion_D2016 <- table_religion_D2016 %>%
  arrange(desc(religion)) %>%
  mutate(lab.ypos1 = cumsum(prop) - 0.5*prop)

table_religion_R2016 <- table_religion_R2016 %>%
  arrange(desc(religion)) %>%
  mutate(lab.ypos2 = cumsum(prop) - 0.5*prop)


ggplot(table_religion_D2016, aes(x ="" , y = prop, fill = religion))+
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start = 0)+
  geom_text(aes(y = lab.ypos1, label = paste(prop,"%", sep = "")), col = "white")+ggtitle("2016 diff religions vote for Democrat")
  
ggplot(table_religion_R2016, aes(x ="" , y = prop, fill = religion))+
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start = 0)+
  geom_text(aes(y = lab.ypos2, label = paste(prop,"%", sep = "")), col = "white")+ggtitle("2016 diff religions vote for Republican")


```
### More male Protestant participate the voting.      

Next, we want to find out what causes Protestant have much proportion on Republican in 2016. From the two bar charts below, we can easily notice that Women tend to vote for Democrat no matter in which religions. If we watch carefully, the male protestant vote for Republican increase from 0.35 percent to almost 0.5 percent. It is possible for Protestant have much proportion on Republican if there are more male Protestants voting and less female Protestants voting.          

```{r echo=FALSE}
#2012
anes_religion_gender_2012 <- anes_use %>% filter(!is.na(vote))%>%filter(!is.na(gender))%>%filter(!is.na(religion))%>%filter(!is.na(church_att)) %>% filter(year=="2012")%>% group_by(gender,religion) %>% count(vote) %>% group_by(gender,religion) %>% mutate(prop=n/sum(n))

ggplot(anes_religion_gender_2012,aes(x=gender,y=prop,fill=vote)) + 
  geom_bar(stat="identity", colour="black")+
  scale_fill_manual(values=c("dodgerblue3", "firebrick2", "gold1", "lightgoldenrod1"))+
  facet_wrap(~religion, ncol=1)+
  labs(title="people in differnt religion voting separate by different gender in 2012 ")

```

```{r echo=FALSE}
#2016
anes_religion_gender_2016 <- anes_use %>% filter(!is.na(vote))%>%filter(!is.na(gender))%>%filter(!is.na(religion))%>%filter(!is.na(church_att)) %>% filter(year=="2016")%>% group_by(gender,religion) %>% count(vote) %>% group_by(gender,religion) %>% mutate(prop=n/sum(n))

ggplot(anes_religion_gender_2016,aes(x=gender,y=prop,fill=vote)) + 
  geom_bar(stat="identity", colour="black")+
  scale_fill_manual(values=c("dodgerblue3", "firebrick2", "gold1", "lightgoldenrod1"))+
  facet_wrap(~religion, ncol=1)+
  labs(title="people in differnt religion voting separate by different gender in 2016 ")

```

### More devout Protestant participate the voting.      

From the bar chart below, Protestant who go to the church every week are tend to vote for Republican and who go to the church less frequently are tend to vote for Democrat. This tendency is not so obvious in 2012 since voters who vote Democrat has a large proportion all the time. However, something change in 2016, Protestant voter who go to the church every week has 65 percent on Republican which approach the maximum and who never go to church has 35 percent on Republican which likely approach the minimum. It is possible for Protestant have much proportion on Republican if there are more devout Protestants (go to the church every week) voting and less false piety.           


```{r echo=FALSE}
#2012
anes_religion_att_2012 <- anes_use %>% filter(!is.na(vote))%>%filter(!is.na(gender))%>%filter(!is.na(religion))%>%filter(!is.na(church_att)) %>% filter(year=="2012")%>% group_by(church_att,religion) %>% count(vote) %>% group_by(church_att,religion) %>% mutate(prop=n/sum(n))

ggplot(anes_religion_att_2012,aes(x=church_att,y=prop,fill=vote)) + 
  geom_bar(stat="identity", colour="black")+
  scale_fill_manual(values=c("dodgerblue3", "firebrick2", "gold1", "lightgoldenrod1"))+
  facet_wrap(~religion, ncol=1)+
  theme(axis.text.x = element_text(angle = 10))+
  labs(title="2012 people in differnt religion voting separate by different church attendence")

```
```{r echo=FALSE}
#2016
anes_religion_att_2016 <- anes_use %>% filter(!is.na(vote))%>%filter(!is.na(gender))%>%filter(!is.na(religion))%>%filter(!is.na(church_att)) %>% filter(year=="2016")%>% group_by(church_att,religion) %>% count(vote) %>% group_by(church_att,religion) %>% mutate(prop=n/sum(n))

ggplot(anes_religion_att_2016,aes(x=church_att,y=prop,fill=vote)) + 
  geom_bar(stat="identity", colour="black")+
  scale_fill_manual(values=c("dodgerblue3", "firebrick2", "gold1", "lightgoldenrod1"))+
  facet_wrap(~religion, ncol=1)+
  theme(axis.text.x = element_text(angle = 10))+
  labs(title="2016 people in differnt religion voting separate by different church attendence")

```

### Democrat candidate makes voter more hopeful and proud and Republican candidate makes voter less hopeful and proud in 2016 than 2012.    

Obviously, the personality of president candidate is an important part in election. Moreover, What candidate let voter feel is an vital part when people voting. From the table below, voters who think Democratic candidate gives them Hopeful and Proud is about 50 percent greater than Republican candidate in 2012. At the same time, voters who think Democratic candidate gives them Angry is also greater than Republican candidate but the result do not change. In 2016, voters who think Democratic candidate gives them Hopeful and Proud still greater than Republican but the difference is not that large.          

Comparing to the line chart above, the decrease of people feel Hopeful and Proud is an significant reason why the reduction exist. On the other hand, although angry and afraid is not unimportant, they affect the voting not as much as the hopeful and proud.         

```{r echo=FALSE}
anes_president=anes_dat%>%
  mutate(
    year=as_factor(VCF0004),
    vote=as_factor(VCF0706),
    Difference=as_factor(VCF0501),
    Democ_Angry=as_factor(VCF0358),
    Democ_Afraid=as_factor(VCF0359),
    Democ_Hopeful=as_factor(VCF0360),
    Democ_Proud=as_factor(VCF0361),
    Rep_Angry=as_factor(VCF0370),
    Rep_Afraid=as_factor(VCF0371),
    Rep_Hopeful=as_factor(VCF0372),
    Rep_Proud=as_factor(VCF0373),
)%>%
filter(year %in% Election_years)

Table_presid<-data.table(anes_president%>%
             select(year,vote,Difference,Democ_Angry,Democ_Afraid,Democ_Hopeful,Democ_Proud,Rep_Angry,Rep_Afraid,Rep_Hopeful,Rep_Proud))%>%filter(!is.na(vote))%>%filter(!is.na(Democ_Angry))%>%filter(!is.na(Democ_Afraid))%>%filter(!is.na(Democ_Hopeful))%>%filter(!is.na(Democ_Proud))%>%filter(!is.na(Rep_Angry))%>%filter(!is.na(Rep_Afraid))%>%filter(!is.na(Rep_Hopeful))%>%filter(!is.na(Rep_Proud))

#For 2016 year
Table_presid_2016<-Table_presid[Table_presid$year=="2016",]



Table_presid_2016$Democ_Angry_count<-ifelse(Table_presid_2016$Democ_Angry=="1. Yes, have felt",1,0)
Table_presid_2016$Democ_Afraid_count<-ifelse(Table_presid_2016$Democ_Afraid=="1. Yes, have felt",1,0)
Table_presid_2016$Democ_Hopeful_count<-ifelse(Table_presid_2016$Democ_Hopeful=="1. Yes, have felt",1,0)
Table_presid_2016$Democ_Proud_count<-ifelse(Table_presid_2016$Democ_Proud=="1. Yes, have felt",1,0)

Table_presid_2016$Rep_Angry_count<-ifelse(Table_presid_2016$Rep_Angry=="1. Yes, have felt",1,0)
Table_presid_2016$Rep_Afraid_count<-ifelse(Table_presid_2016$Rep_Afraid=="1. Yes, have felt",1,0)
Table_presid_2016$Rep_Hopeful_count<-ifelse(Table_presid_2016$Rep_Hopeful=="1. Yes, have felt",1,0)
Table_presid_2016$Rep_Proud_count<-ifelse(Table_presid_2016$Rep_Proud=="1. Yes, have felt",1,0)

num_Democ_Angry2<-sum(Table_presid_2016$Democ_Angry_count)
num_Democ_Afraid2<-sum(Table_presid_2016$Democ_Afraid_count)
num_Democ_Hopeful2<-sum(Table_presid_2016$Democ_Hopeful_count)
num_Democ_Proud2<-sum(Table_presid_2016$Democ_Proud_count)

num_Rep_Angry2<-sum(Table_presid_2016$Rep_Angry_count)
num_Rep_Afraid2<-sum(Table_presid_2016$Rep_Afraid_count)
num_Rep_Hopeful2<-sum(Table_presid_2016$Rep_Hopeful_count)
num_Rep_Proud2<-sum(Table_presid_2016$Rep_Proud_count)

#For 2012 year
Table_presid_2012<-Table_presid[Table_presid$year=="2012",]



Table_presid_2012$Democ_Angry_count<-ifelse(Table_presid_2012$Democ_Angry=="1. Yes, have felt",1,0)
Table_presid_2012$Democ_Afraid_count<-ifelse(Table_presid_2012$Democ_Afraid=="1. Yes, have felt",1,0)
Table_presid_2012$Democ_Hopeful_count<-ifelse(Table_presid_2012$Democ_Hopeful=="1. Yes, have felt",1,0)
Table_presid_2012$Democ_Proud_count<-ifelse(Table_presid_2012$Democ_Proud=="1. Yes, have felt",1,0)

Table_presid_2012$Rep_Angry_count<-ifelse(Table_presid_2012$Rep_Angry=="1. Yes, have felt",1,0)
Table_presid_2012$Rep_Afraid_count<-ifelse(Table_presid_2012$Rep_Afraid=="1. Yes, have felt",1,0)
Table_presid_2012$Rep_Hopeful_count<-ifelse(Table_presid_2012$Rep_Hopeful=="1. Yes, have felt",1,0)
Table_presid_2012$Rep_Proud_count<-ifelse(Table_presid_2012$Rep_Proud=="1. Yes, have felt",1,0)

num_Democ_Angry1<-sum(Table_presid_2012$Democ_Angry_count)
num_Democ_Afraid1<-sum(Table_presid_2012$Democ_Afraid_count)
num_Democ_Hopeful1<-sum(Table_presid_2012$Democ_Hopeful_count)
num_Democ_Proud1<-sum(Table_presid_2012$Democ_Proud_count)

num_Rep_Angry1<-sum(Table_presid_2012$Rep_Angry_count)
num_Rep_Afraid1<-sum(Table_presid_2012$Rep_Afraid_count)
num_Rep_Hopeful1<-sum(Table_presid_2012$Rep_Hopeful_count)
num_Rep_Proud1<-sum(Table_presid_2012$Rep_Proud_count)


#2012 bar chart
bp_2012<-data.frame(party=rep(c("Democratic","Republican"),each=4),
               affect=rep(c("Angry","Afraid","Hopeful","Proud"),2),
               count=c(num_Democ_Angry1,num_Democ_Afraid1,num_Democ_Hopeful1,num_Democ_Proud1,num_Rep_Angry1,num_Rep_Afraid1,num_Rep_Hopeful1,num_Rep_Proud1))

#2016 bar chart
bp_2016<-data.frame(party=rep(c("Democratic","Republican"),each=4),
               affect=rep(c("Angry","Afraid","Hopeful","Proud"),2),
               count=c(num_Democ_Angry2,num_Democ_Afraid2,num_Democ_Hopeful2,num_Democ_Proud2,num_Rep_Angry2,num_Rep_Afraid2,num_Rep_Hopeful2,num_Rep_Proud2))

#plot the bar chart

ggplot(data=bp_2012, aes(x=affect, y=count, fill=party)) +
  geom_bar(stat="identity", position="dodge2")+
  ggtitle("2012 Democratic and Republican Presidential Cand affect barchart")+
  geom_text(aes(label=count), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  scale_fill_brewer(palette="Paired")+theme_minimal()


ggplot(data=bp_2016, aes(x=affect, y=count, fill=party)) +
  geom_bar(stat="identity", position="dodge2")+
  ggtitle("2016 Democratic and Republican Presidential Cand affect barchart")+
  geom_text(aes(label=count), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  scale_fill_brewer(palette="Set2")+theme_minimal()

```


```{r echo=FALSE}
bp_total<-data.frame(year=rep(c("2012","2016"),each=8),
                     party=rep(c("Democratic","Republican","Democratic","Republican"),each=4),
               affect=rep(c("Angry","Afraid","Hopeful","Proud"),2),
               count=c(num_Democ_Angry1,num_Democ_Afraid1,num_Democ_Hopeful1,num_Democ_Proud1,num_Rep_Angry1,num_Rep_Afraid1,num_Rep_Hopeful1,num_Rep_Proud1,num_Democ_Angry2,num_Democ_Afraid2,num_Democ_Hopeful2,num_Democ_Proud2,num_Rep_Angry2,num_Rep_Afraid2,num_Rep_Hopeful2,num_Rep_Proud2))




datatable(bp_total, class = 'cell-border stripe')
```




On conclusion, there are some reasons why a huge reduction exist. The first reason is there are more male Protestants participate the voting since the male protestant more tend to vote for Republican. The second reason is there are more devout Protestant participate the voting since the Protestants who go to the church more often are more tend to vote for Republican. The third reason is that Democrat candidate in 2016 are less hopeful and proud than 2012 and Republican candidate in 2016 are more hopeful and proud than 2012.       





















